---
title: "Data Wrangling"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

#  Let's import the necessary libraries.

```{r echo=FALSE}
#install.packages("tidyverse")
library("tidyverse")
```

# *1) Let's import the datasets and inspect them*

```{r}
ideal1<- read.csv("https://raw.githubusercontent.com/ThumbiMwangi/R_sources/master/ideal1.csv")
```

```{r}
ideal2<- read.csv("https://raw.githubusercontent.com/ThumbiMwangi/R_sources/master/ideal2.csv")
```



```{r}
#getwd()
```


```{r}
#write.csv(ideal1, "ideal1.csv")
```


```{r}
#write.csv(ideal2, "ideal2.csv")
```


```{r}
ideal1 %>% dim()
```


```{r}
ideal2 %>% dim()
```
```{r}
ideal1 %>% colnames()
```
```{r}
ideal2 %>% colnames()
```


## Ideal1 has 30 lines and 8 columns whereas ideal2 has 270 lines and 9 columns.


# *2) Changing the CADOB column in ideal 1 into a date and the VisitDate column into ideal 2 into a date.*

```{r}
ideal1$CADOB %>% typeof()
```

```{r}
ideal2$VisitDate %>% typeof()
```


```{r}
ideal1$CADOB<- as.Date(ideal1$CADOB)
```


```{r}
ideal2$VisitDate <- as.Date(ideal2$VisitDate)
```

```{r}
ideal1$CADOB %>% typeof()
```


```{r}
ideal2$VisitDate %>% typeof()
```
## The missing values and duplicates in each dataset


```{r}
ideal1 %>% is.na() %>% colSums()
```
### ***The ReasonsLoss has 21 missing values***

```{r}
ideal2 %>% is.na() %>% colSums()
```

```{r}
ideal1 %>% duplicated() %>% sum()
```

```{r}
ideal2 %>% duplicated() %>% sum()
```


# *3) Recode the CalfSex column into "1"="Male", "2"="Female" and the ReasonsLoss column into: 6= «owner declined»,  7= «Death», 12= «Relocation», 18= «Given out» as well.*


```{r}
ideal1 <-ideal1 %>% mutate( CalfSex1=
  case_when(
   CalfSex=="1" ~ "Male", 
   CalfSex=="2" ~ "Female"
  )
)
```

```{r}
ideal1$ReasonsLoss %>% unique()
```
```{r}
ideal1$ReasonsLoss %>% class()
```



```{r}
ideal1 %>% mutate( ReasonsLoss1=
  case_when(
   ReasonsLoss==6 ~ "owner declined", 
   ReasonsLoss==7 ~ "Death",
   ReasonsLoss==12~ "Relocation",
   ReasonsLoss==18 ~ "Given out"
  )
)
```


# *4) Order the ideal1 dataset by the CADOB column in descending order*

```{r}
ideal1$CalfID %>% unique %>% length == dim(ideal1)[1]
```



```{r}
ideal1<-ideal1 %>% group_by(CalfID) %>% arrange(desc(CADOB))
```



# *5) Grouping the ideal1 dataset by CalfID and creating a new column minweight as minweight=min(RecruitWeigh) and Puting the result in a new dataset named ideal1a*


```{r}
ideal1a<-ideal1 %>% group_by(CalfID) %>% 
  mutate(minweight=min(RecruitWeight, na.rm = T))
```



# *6) Let's full Join the two datasets ideal1a and ideal2 by CalfID  and call it ideal3*

```{r}
ideal3 <- ideal1a %>% 
  full_join(ideal2, by="CalfID")
```


```{r}
ideal1a %>% colnames()
```
```{r}
ideal2 %>% colnames()
```

```{r}
ideal3 %>% colnames()
```

```{r}
ideal3 %>% is.na() %>% colSums()
```


# 7) Let's extract visit_number from the 4th to the 5th character of visitID


```{r}
ideal3 <- ideal3 %>% 
  mutate(visit_number= str_sub(VisitID,start=4,end=5))
```


```{r}
ideal3 %>% head()
```

# 8) Counting the number of sublocation per visit_number in ideal3

```{r}
ideal3 %>% group_by(visit_number) %>% count(sublocation)
```

# *9) Plot the survival of calves in the first year of life*

```{r}
calves_recruited<- ideal3 %>%  group_by(sublocation) %>% summarise(freq=n())
```



```{r}
ggplot(calves_recruited, aes(y=sublocation, x=freq))+
  geom_col()+
  labs(y="sublocation", x="frequency")
```



# *10) Plot the number of calves recruited in each sublocation in descending order.*

```{r}
ggplot(calves_recruited, aes(y=reorder(sublocation,freq), x=freq))+
  geom_col()+
  labs(y="sublocation", x="frequency")
```

# *11) Plot the number of calves recruited in each sublocation by gender in descending order.*

```{r}
calves_recruited1<- ideal3 %>%  group_by(sublocation,CalfSex1) %>% summarise(freq=n())
```


```{r}
ggplot(calves_recruited1, aes(y=reorder(sublocation,freq), x=freq, fill=  CalfSex1))+
  geom_col()+
  labs(y="sublocation", x="frequency")
```


```{r}
ideal3 %>% summary()
```



# *13) Droppind the former sex variable and renaming the new one.*


```{r}
ideal3 <-ideal3 %>% select(-CalfSex) %>% 
  rename(CalfSex=CalfSex1)
```



# *14) Subset the male dataset and name it ideal3_m*

```{r}
ideal3_m <-ideal3 %>% filter(CalfSex=="Male")
```

# *15) Plot a boxplot of RecruitWeight*


```{r}
ggplot(ideal3_m, aes(y=RecruitWeight))+
  geom_boxplot()
  #+ labs(y="sublocation", x="frequency")
```

# *16) Plot a boxplot of RecruitWeight per Sex*

```{r}
ggplot(ideal3, aes(y=RecruitWeight, fill=  CalfSex))+
  geom_boxplot()
  #+ labs(y="sublocation", x="frequency")
```

```{r}
ggplot(ideal3, aes(x=RecruitWeight))+
  geom_histogram()
  #+ labs(y="sublocation", x="frequency")
```

```{r}
knitr::opts_chunk$set(echo = TRUE, fig.width = 40, fig.height = 10)
```


# *17)Plot the scatterplot of  RecruitWeight and Weight*

```{r}
ggplot(ideal3, aes(x=RecruitWeight, y= Weight))+
  geom_point()
  #+ labs(y="sublocation", x="frequency")
```

